Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-05-03 inc/extension/external_forces/formal.tex
fixed some operator/matrix symbology

For constant static forces having no dependence in $ x $, $ y $, $ z $, or $ t $, one can readily show that there is no effect on the pulse propagation parameters $ \sigma_{ \alpha } $, $ \eta_{ \alpha } $, and $ \gamma_{ \alpha } $.
For example, consider an acceleration electric field $ \vec{E} = -E \hat{z} $ such that electrons are accelerated in the positive $ z $ direction.
The force is then $ \vec{F}_{ext} = e E \hat{z} $, so that from \ref{eq:Kint} we have
\begin{equation}
  K^{ext}_{ij} = \frac{e E} {N} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x d^{3}p } \,\text{.}
\end{equation}
It is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ (or equivalently $ K^{ext} = \hat{0} $) meaning that this static field has no effect on the parameters of the pulse.
Of course, this is only true in the pulse frame, the frame of the simulation, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m $ is unchanged.
In the lab frame however, where  $ \Delta E = p \Delta p / m $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.

@@

For the case of a cylindrically symmetric perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse: 
\begin{equation} \label{eq:linear_force}
\vec{F}_{ext} = - M_{ \smallT } r \hat{\operatorname{r}} - M_{z} z \hat{\operatorname{z}} \,\text{.}
\end{equation}
Here, $M_{ \smallT }$ and $M_{z}$ characterize the strength of a lens or a compressor, respectively, and are chosen to be positive for focusing/compression elements.
Substitution into the integral equation for the $ K^{ext} $ matrix elements,
\begin{equation}
  \begin{split} 
    K^{ext}_{ij} & \equiv \dfrac{1} {N} \int u_{i} u_{j} ( \vec{\nabla}_{p} f ) \cdot \vec{F}_{ext} d^{3} x d^{3} p \\ 
    & = \dfrac{1} {N} \int u_{i} u_{j} \biggl[ x \left( \dfrac{ p_{x} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } x } { \sigma_{ \smallT } \eta_{ \smallT } } \right) + y \left( \dfrac{ p_{y} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } y } { \sigma_{ \smallT } \eta_{ \smallT } } \right) \\
    & \quad + z \left( \dfrac{ p_{z} } { \eta_{z} } - \dfrac{ \gamma_{z} z } { \sigma_{z} \eta_{z} } \right) \biggr] f d^{3} x d^{3} p \,\text{,}
  \end{split}
\end{equation}
leads to the $6\times6$ matrix
\begin{equation}
  K^{ext} = 
  \begin{pmatrix}
    \hat{k}^{lin}_{\smallT} & \hat{0} & \hat{0} \\
    \hat{0} & \hat{k}^{lin}_{\smallT} & \hat{0} \\
    \hat{0} & \hat{0} & \hat{k}^{lin}_{z}
  \end{pmatrix} \text{,}
\end{equation}
where
\begin{equation}
  \hat{k}^{lin}_{\alpha} =
  \begin{pmatrix}
    0 & M_{\alpha} \sigma_{\alpha} \\
    M_{\alpha} \sigma_{\alpha} & 2 M_{\alpha} \gamma_{\alpha}
  \end{pmatrix} \text{.}
\end{equation}
which by \ref{eq:dainvdt} this yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
This derivation has been for a single force, in practice however there may be multiple forces.
Therefore if each force is labeled by $f$, \ref{eq:ag_big_gamma_gamma} may be restated as
\begin{equation}
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} \,\text{.}
\end{equation}

>>>>

For constant static forces having no dependence in $ x $, $ y $, $ z $, or $ t $, one can readily show that there is no effect on the pulse propagation parameters $ \sigma_{ \alpha } $, $ \eta_{ \alpha } $, and $ \gamma_{ \alpha } $.
For example, consider an acceleration electric field $ \vec{E} = -E \hat{z} $ such that electrons are accelerated in the positive $ z $ direction.
The force is then $ \vec{F}_{ext} = e E \hat{z} $, so that from \ref{eq:Kint} we have
\begin{equation}
  K^{ext}_{ij} = \frac{e E} {N} \int{ u_{i} u_{j} \left( \frac{p_{z}} {\eta_{z}} - \frac{\gamma_{z} z} {\sigma_{z} \eta_{z}} \right) f d^{3}x d^{3}p } \,\text{.}
\end{equation}
It is clear that no function of this form multiplying $ f $, which is a Gaussian distribution, can avoid creating an odd integrand, and thus $ K_{ij}^{ext} = 0 $ for all $ i $ and $ j $ meaning that this static field has no effect on the parameters of the pulse.
Of course, this is only true in the pulse frame, the frame of the simulation, where there is no total momentum and thus $ \Delta E = \left( \Delta p \right)^{2} / 2 m $ is unchanged.
In the lab frame however, where  $ \Delta E = p \Delta p / m $, there is an increase in the momentum spread, as the pulse now has the large momentum $p$ imparted by the accelerating field.

@@

For the case of a cylindrically symmetric perfect lens, or in the longitudinal direction a perfect RF compression element, with a parabolic potential, the external force is linear with distance from the center of the electron pulse: 
\begin{equation} \label{eq:linear_force}
\vec{F}_{ext} = - M_{ \smallT } r \hat{r} - M_{z} z \hat{z} \,\text{.}
\end{equation}
Here, $M_{ \smallT }$ and $M_{z}$ characterize the strength of a lens or a compressor, respectively, and are chosen to be positive for focusing/compression elements.
Substitution into the integral equation for the $ K^{ext} $ matrix elements,
\begin{equation}
  \begin{split} 
    K^{ext}_{ij} & \equiv \dfrac{1} {N} \int u_{i} u_{j} ( \vec{\nabla}_{p} f ) \cdot \vec{F}_{ext} d^{3} x d^{3} p \\ 
    & = \dfrac{1} {N} \int u_{i} u_{j} \biggl[ x \left( \dfrac{ p_{x} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } x } { \sigma_{ \smallT } \eta_{ \smallT } } \right) + y \left( \dfrac{ p_{y} } { \eta_{ \smallT } } - \dfrac{ \gamma_{ \smallT } y } { \sigma_{ \smallT } \eta_{ \smallT } } \right) \\
    & \quad + z \left( \dfrac{ p_{z} } { \eta_{z} } - \dfrac{ \gamma_{z} z } { \sigma_{z} \eta_{z} } \right) \biggr] f d^{3} x d^{3} p \,\text{,}
  \end{split}
\end{equation}
leads to the $6\times6$ matrix
\begin{equation}
  K^{ext} = 
  \begin{pmatrix}
    k^{lin}_{\smallT} & 0 & 0 \\
    0 & k^{lin}_{\smallT} & 0 \\
    0 & 0 & k^{lin}_{z}
  \end{pmatrix} \text{,}
\end{equation}
where
\begin{equation}
  k^{lin}_{\alpha} =
  \begin{pmatrix}
    0 & M_{\alpha} \sigma_{\alpha} \\
    M_{\alpha} \sigma_{\alpha} & 2 M_{\alpha} \gamma_{\alpha}
  \end{pmatrix} \text{.}
\end{equation}
which by \ref{eq:dainvdt} this yields an additional term $M_{\alpha} \sigma_{\alpha}$ which is added to the original $\dfrac{d \gamma_{\alpha}}{d t}$ equation.
This derivation has been for a single force, in practice however there may be multiple forces.
Therefore if each force is labeled by $f$, \ref{eq:ag_big_gamma_gamma} may be restated as
\begin{equation}
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} \,\text{.}
\end{equation}
